---
title: "nicolaus"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Provides access to, and download from, [Copernicus Marine Data Store's catalog](https://marine.copernicus.eu/) using R language. 

# Copernicus resources

Copernicus serves **so many** data resources; finding what you want can be a challenge.  Check out the new [Marine Data Store](https://marine.copernicus.eu/news/introducing-new-copernicus-marine-data-store). And checkout the listing [here](https://marine.copernicus.eu/about/producers).

## Data catalogs

Like data offerings from [OBPG](https://oceancolor.gsfc.nasa.gov/), Copernicus strives to provided consistent dataset identifiers that are easily decoded programmatically (and with practice by eye). In order to download programmatically you must have the `product_id` and or `dataset_id` in hand. Learn more about Copernicus [nomenclature rules here](https://help.marine.copernicus.eu/en/articles/6820094-how-is-defined-the-nomenclature-of-copernicus-marine-data#h_34a5a6f21d).  

## `get` or `subset`

The [Copernicus Marine Toolbox](https://help.marine.copernicus.eu/en/collections/4060068-copernicus-marine-toolbox) command-line application, `copernicus-marine` provides two primary methods for donwloading data: `get` and `subset`. `get` is not well documented, but subset does what it implies - subsetting resources by variable, spatial bounding box, depth and time.  This package only supports `subset`.

## Requirements

  + [R v4.1+](https://www.r-project.org/)
  
### From CRAN
  + [rlang](https://CRAN.R-project.org/package=rlang)
  + [dplyr](https://CRAN.R-project.org/package=dplyr)
  + [readr](https://CRAN.R-project.org/package=readr)
  + [jsonlite](https://CRAN.R-project.org/package=jsonlite)
  
  
### From github
  + [copernicus](https://github.com/BigelowLab/copernicus)
  

## Installation

```
remotes::install_github("BigelowLab/nicolaus")
```

## Configuration

If you have already installed and configured the [copernicus R package](https://github.com/BigelowLab/copernicus) then you are all set to use this package.  If you have not installed and configured the [copernicus R package](https://github.com/BigelowLab/copernicus) then please do that first.


## Fetch the products catalog

You can download a product catalog for local storage.

```{r}
suppressPackageStartupMessages({
  library(nicolaus)
  library(dplyr)
})

ok = fetch_catalog()
ok
```


## Import the products catalog

This downloads into a "catalogs" directory within your [copernicus](https://github.com/BigelowLab/copernicus) data directory Now read it in.

```{r read_catalog}
x = import_catalog()
dplyr::glimpse(x)
```

By default this provides a flattened table of available products and their datasets along with each dataset's variables. There is a lot of information in the JSON catalog that we place in the table including:

 + depth min, max and depth layer count
 
 + time min, max and step size
 
 + short and standard name for each variable
 

## Save the products catalog as a table

It doesn't take a long time to parse the JSON list into a table, but it is worth the effort to save the table as a flat structure for later us.  

```{r}
x = write_catalog(x)
y = read_catalog()
identical(x,y)
```

